The Lipid Interactome Repository

### 2025-06-16 Background/MultifunctionalLipidProbesOverview.qmd
Finally getting the bf-GPI data page online
Updating with data wrangling scripts and some suggested reviewer changes

Functionalized lipid probes are a toolset intended to fuse several niche biotechnologies. This overview will serve as a summary of the capabilities, nuances, and limitations of their utility. Please find additional resources and published primary and review articles below for greater detail.

@@

Among the various crosslinking motifs, the diazirine ring is the smallest, making it the least likely to interfere with a lipid’s natural interactions. Still, the introduction of diazirine into a fatty acid side chain will slightly decrease hydrophobicity and lipid ordering. Upon exposure to 355 nm UV light, the diazirine ring forms a reactive intermediate, which rapidly form covalent bonds with nearby molecules. This reaction is highly versatile, as reactive intermediates such as carbenes can insert into nearby carbon-hydrogen, sulfur-hydrogen, nitrogen-hydrogen, oxygen-hydrogen, or double bonds, allowing for broad coverage of potential interaction partners1. Despite these numerous options, reaction with glutamate seems to be modestly preferred2.
The small size and high reactivity of diazirine crosslinking make it an invaluable tool for capturing lipid-protein interactions while minimizing perturbations to native lipid function.

>>>>

Functionalized lipid probes are a toolset intended to fuse several niche biotechnologies. This overview will serve as a summary of the capabilifties, nuances, and limitations of their utility. Please find additional resources and published primary and review articles below for greater detail.

@@

Among the various crosslinking motifs, the diazirine ring is the smallest, making it the least likely to interfere with a lipid’s natural interactions. Still, the introduction of diazirine into a fatty acid side chain will slightly decrease hydrophobicity and lipid ordering. Upon exposure to 355 nm UV light, the diazirine ring forms a reactive intermediate, which rapidly form covalent bonds with nearby molecules. This reaction is highly versatile, as reactive intermediates such as carbenes can insert into nearby carbon-hydrogen, sulfur-hydrogen, nitrogen-hydrogen, oxygen-hydrogen, or double bonds, allowing for broad coverage of potential interaction partners[@haberkantFatFabulousBifunctional2014]. Despite these numerous options, reaction with glutamate seems to be modestly preferred[@westDesignEvaluationCyclobutane2022].
The small size and high reactivity of diazirine crosslinking make it an invaluable tool for capturing lipid-protein interactions while minimizing perturbations to native lipid function.

### 2025-06-16 IndividualStudies/DCC_2025.qmd
first update in a long time, minor changes
Updating with data wrangling scripts and some suggested reviewer changes

```{r HiddenOriginalDataWrangling, include=FALSE, echo=FALSE}

# data <- readr::read_csv(paste0(here::here(), "/IndividualStudies/DataTables/DCC_2025_ChemrXiv_FullDataset.csv")) |> 
#   group_by(gene_name) |>
#   filter(ratio_MvsL != "") |>
#   summarise(logFC = mean(log2(ratio_MvsL)),
#             pvalue = mean(pvalue),
#             numReplicates = n(),
#             AveExpr = log10(mean(L_abundance) + mean(M_abundance))/2) |>
#   arrange(-logFC) |>
#   mutate(LipidProbe = "BF-NAPE",
#          hit_annotation = case_when(.default = "no hit",
#                                     logFC >= 0.5849625 & pvalue <= 0.05 & numReplicates == 2 ~ "enriched candidate",
#                                     logFC >= 0.5849625 & pvalue <= 0.05 & numReplicates == 3 ~ "enriched hit")) |>
#   glimpse()

# median_logFC <- median(data$logFC)

# data$logFC <- data$logFC - median_logFC

@@

```{r, echo=FALSE, warning=FALSE}
#| fig-cap: "Volcano plots depict the relative enrichment of each protein versus non-crosslinked control (x-axis) and the statistical significance of each protein following a Student's t-test (y-axis, -log10 transformed). Black proteins are identified in only one replicate, Purple enriched candidates are defined as proteins identified in two replicates & fold-change > 1.5 (logFC 0.585) & pvalue < 0.01, and Orange enriched hits are defined as proteins identified in all three replicates & fold-change > 1.5 (logFC 0.585) & pvalue < 0.01)."

@@

```{r, echo=FALSE, warning=FALSE}
#| fig-cap: "Ranked-order plots depict the relative enrichment of each protein versus non-crosslinked control (y-axis) from lowest to highest. Black proteins are identified in only one replicate, Purple enriched candidates are defined as proteins identified in two replicates & fold-change > 1.5 (logFC 0.585) & pvalue < 0.05, and Orange enriched hits are defined as proteins identified in all three replicates & fold-change > 1.5 (logFC 0.585) & pvalue < 0.05)."

@@

```{r, echo=FALSE, warning=FALSE}
#| fig-cap: "MA plots depict the average abundance of each protein in the crosslinked and uncrosslinked conditions (x-axis) versus the log2-transformed fold-change between the crosslinked and uncrosslinked conditions (y-axis). Black proteins are identified in only one replicate, Purple enriched candidates are defined as proteins identified in two replicates & fold-change > 1.5 (logFC 0.585) & pvalue < 0.05, and Orange enriched hits are defined as proteins identified in all three replicates & fold-change > 1.5 (logFC 0.585) & pvalue < 0.05)."

@@

###### Wrangling for the Venn Diagram ######
# Read the dataset
VennData <- readr::read_csv(paste0(here::here(), "/IndividualStudies/DataTables/DCC_2025_ChemrXiv_FullDataset.csv")) |>

  filter(ratio_MvsL > log2(1.5), pvalue < 0.05) |>
  glimpse()

@@

data <- data |>
  dplyr::select(gene_name, LipidProbe, hit_annotation, logFC, pvalue, numReplicates) |>
  mutate(logFC = signif(logFC, digits=3),
          pvalue = signif(pvalue, digits=3)) |>
  arrange(hit_annotation, numReplicates, -logFC)

@@

## Displays the reactive dataframe
datatable(data_reactive, 
          caption = "This datatable contains all the data plotted above. Check the boxes below to filter the dataset by which lipid probe was used in the pulldown and by significance thresholds. logFC ratios are median normalized. \"Num Replicate IDs\" denotes the number of replicates in which the protein was identified. As above, \"enriched hit\" denotes proteins identified in all three replicates & fold-change > 1.5 (logFC 0.585) & pvalue < 0.05. \"enriched candidates\" are proteins identified in two replicates & fold-change > 1.5 (logFC 0.585) & pvalue < 0.05.",
          options = list(pageLength = 10),
                         colnames = c("Gene Name", "Lipid Probe", "Hit Annotation", "LogFC", "p value", "Num Replicate IDs"),
                         rownames=FALSE)

>>>>

@@

```{r Volcanoplots, echo=FALSE, warning=FALSE}
#| fig-cap: "Volcano plots depict the relative enrichment of each protein versus non-crosslinked control (x-axis) and the statistical significance of each protein following a Student's t-test (y-axis, -log10 transformed). Black proteins are identified in only one replicate, Purple enriched candidates are defined as proteins identified in two replicates & fold-change > 1.5 (logFC 0.585) & pvalue < 0.01, and Orange enriched hits are defined as proteins identified in all three replicates & fold-change > 1.5 (logFC 0.585) & pvalue < 0.01)."

@@

```{r ROplots, echo=FALSE, warning=FALSE}
#| fig-cap: "Ranked-order plots depict the relative enrichment of each protein versus non-crosslinked control (y-axis) from lowest to highest. Black proteins are identified in only one replicate, Purple enriched candidates are defined as proteins identified in two replicates & fold-change > 1.5 (logFC 0.585) & pvalue < 0.05, and Orange enriched hits are defined as proteins identified in all three replicates & fold-change > 1.5 (logFC 0.585) & pvalue < 0.05)."

@@

```{r MAplots, echo=FALSE, warning=FALSE}
#| fig-cap: "MA plots depict the average abundance of each protein in the crosslinked and uncrosslinked conditions (x-axis) versus the log2-transformed fold-change between the crosslinked and uncrosslinked conditions (y-axis). Black proteins are identified in only one replicate, Purple enriched candidates are defined as proteins identified in two replicates & fold-change > 1.5 (logFC 0.585) & pvalue < 0.05, and Orange enriched hits are defined as proteins identified in all three replicates & fold-change > 1.5 (logFC 0.585) & pvalue < 0.05)."

@@

###### Wrangling for the Venn Diagram ######
# Read the dataset
VennData <- readr::read_csv(here("IndividualStudies/DataTables/DataWrangling/DCC_2025_ChemrXiv_FullDataset.csv")) |>
  filter(ratio_MvsL > log2(1.5), pvalue < 0.05) |>
  glimpse()

@@

data <- data |>
  dplyr::select(gene_name, gene_ID, LipidProbe, hit_annotation, logFC, pvalue, numReplicates) |>
  mutate(logFC = signif(logFC, digits=3),
          pvalue = signif(pvalue, digits=3)) |>
  arrange(hit_annotation, numReplicates, -logFC)

@@

## Displays the reactive dataframe
datatable(data_reactive, 
          caption = "This datatable contains all the data plotted above. Check the boxes below to filter the dataset by which lipid probe was used in the pulldown and by significance thresholds. logFC ratios are median normalized. \"Num Replicate IDs\" denotes the number of replicates in which the protein was identified. As above, \"enriched hit\" denotes proteins identified in all three replicates & fold-change > 1.5 (logFC 0.585) & pvalue < 0.05. \"enriched candidates\" are proteins identified in two replicates & fold-change > 1.5 (logFC 0.585) & pvalue < 0.05.",
          options = list(pageLength = 10),
                         colnames = c("Gene Name", "Gene ID/Accession", "Lipid Probe", "Hit Annotation", "LogFC", "p value", "Num Replicate IDs"),
                         rownames=FALSE)

### 2025-06-16 IndividualStudies/SK_2024.qmd
Updating with data wrangling scripts and some suggested reviewer changes

```{r AddingToSearchByProteinPage, include=FALSE, echo=FALSE}

# data <- readr::read_csv(here::here("IndividualStudies/DataTables/SK_2024_ChemEuro.csv")) |>
#   dplyr::select(LipidProbe, gene_name, logFC, pvalue, hit_annotation) |>
#   glimpse()

# CombinedProbeData <- read_csv(here("ShinyApps/LipidInteractomics_ShinyApp/combinedProbeDatasets_TMT.csv")) |>
#   dplyr::select(-`...1`) |>
#   glimpse() |>
#   rbind(data)

# write_csv(CombinedProbeData, here("ShinyApps/LipidInteractomics_ShinyApp/combinedProbeDatasets_TMT.csv"))

```{r HiddenOriginalDataWrangling, include=FALSE, echo=FALSE}

# gene_names <- read_csv(here("IndividualStudies/DataTables/gene_symbol_results.csv")) |>
#   mutate(
#     gene_name = case_when(
#       gene_symbol == "unnamed protein product [Homo sapiens]" ~ accession,
#       gene_symbol == "unknown [Homo sapiens]" ~ accession,
#       TRUE ~ str_remove(gene_symbol, "\"$")
#     ) 
#   ) |>
#   mutate(Accession = accession) |>
#   dplyr::select(-gene_symbol, -accession) |>
#   glimpse()

# rawData <- readr::read_csv(here("IndividualStudies/DataTables/SK_2024_fullDataset_originalHandoff.csv")) |>
#   dplyr::select(Accession, Description, `Abundance Ratio (log2): (Penta) / (Control)`, `Abundance Ratio Adj. P-Value: (Penta) / (Control)`, `Abundances (Grouped): Control`, `Abundances (Grouped): Penta`) |>
#   rename(logFC = `Abundance Ratio (log2): (Penta) / (Control)`,
#           pvalue = `Abundance Ratio Adj. P-Value: (Penta) / (Control)`,
#          Control_abundance = `Abundances (Grouped): Control`,
#          Penta_abundance = `Abundances (Grouped): Penta`) |>
#   mutate(AveExpr = log2((Control_abundance + Penta_abundance)/2)) |>
#   left_join(gene_names, by = "Accession") |>
#   dplyr::select(gene_name, Accession, Description, logFC, pvalue, AveExpr, Control_abundance, Penta_abundance) |>
#   mutate(hit_annotation = case_when(.default = "no hit",
#                                     logFC >= 2 & pvalue <= 0.05 ~ "enriched hit",
#                                     logFC >= 1 & pvalue < 0.05 ~ "enriched candidate")) |>
#   mutate(LipidProbe = "bf-GPI") |>
#   glimpse()

# write_csv(rawData, here("IndividualStudies/DataTables/SK_2024_ChemEuro.csv"))

# library(biomaRt)

# # Try connecting to a mirror
# ensembl <- useEnsembl(biomart = "ensembl", 
#                       dataset = "hsapiens_gene_ensembl",
#                       mirror = "useast")  # or "uswest", "asia"

# # Function to get gene symbol or fallback gene title
# get_gene_symbol <- function(accession) {
#   tryCatch({
#     search <- entrez_search(db = "protein", term = accession)
#     if (length(search$ids) == 0) return(NA)

#     gb_record <- entrez_fetch(db = "protein", id = search$ids[[1]], rettype = "gb", retmode = "text")
#     match <- regmatches(gb_record, regexpr("/gene=\"[^\"]+\"", gb_record))

#     gene <- if (length(match) > 0) {
#       sub("/gene=\"|\"", "", match)
#     } else {
#       summary <- entrez_summary(db = "protein", id = search$ids[[1]])
#       summary$title
#     }

#     return(gene)
#   }, error = function(e) {
#     message("Error with accession ", accession, ": ", e$message)
#     return(NA)
#   })
# }

# # Your vector of accession numbers
# accessions <- pull(rawData, Accession)

# # Path to your CSV results file
# csv_file <- here("IndividualStudies/DataTables/gene_symbol_results.csv")

# # Load previously saved results if the file exists
# if (file.exists(csv_file)) {
#   results <- read.csv(csv_file, stringsAsFactors = FALSE)
#   processed <- results$accession
# } else {
#   results <- data.frame(accession = character(), gene_symbol = character(), stringsAsFactors = FALSE)
#   processed <- character()
# }

# # Iterate through accessions, skipping ones already done
# for (acc in accessions) {
#   if (acc %in% processed) next  # skip if already processed

#   gene <- get_gene_symbol(acc)
#   results <- rbind(results, data.frame(accession = acc, gene_symbol = gene, stringsAsFactors = FALSE))

#   # Save progress after every accession
#   write.csv(results, csv_file, row.names = FALSE)

#   message("Processed: ", acc)
# }

@@

data <- data |>
  dplyr::select(gene_name, LipidProbe, hit_annotation, logFC, pvalue) |>
  mutate(logFC = signif(logFC, digits=3),
          pvalue = signif(pvalue, digits=3)) |>
  arrange(hit_annotation) |>
  arrange(LipidProbe)

@@

## Displays the reactive dataframe
datatable(data_reactive, 
          caption = "'No hit' proteins are unenriched or depleted in the presence of probe, 'enriched candidates' are defined as proteins with a p-value < 0.05 and a fold-change > 2 (log2FC > 1) for the bf-GPI versus control probe, and 'enriched hits' are defined as proteins with a p-value < 0.05 and a fold change of at least 4-fold (log2FC >2) for the bf-GPI versus control probe", 
          options = list(pageLength = 15),
                         colnames = c("Gene Name", "Lipid Probe", "Hit Annotation", "LogFC", "p value"))

>>>>

@@

data <- data |>
  dplyr::select(gene_name, Accession, LipidProbe, hit_annotation, logFC, pvalue) |>
  mutate(logFC = signif(logFC, digits=3),
          pvalue = signif(pvalue, digits=3)) |>
  arrange(hit_annotation) |>
  arrange(LipidProbe)

@@

## Displays the reactive dataframe
datatable(data_reactive, 
          caption = "'No hit' proteins are unenriched or depleted in the presence of probe, 'enriched candidates' are defined as proteins with a p-value < 0.05 and a fold-change > 2 (log2FC > 1) for the bf-GPI versus control probe, and 'enriched hits' are defined as proteins with a p-value < 0.05 and a fold change of at least 4-fold (log2FC >2) for the bf-GPI versus control probe", 
          options = list(pageLength = 15),
                         colnames = c("Gene Name", "Gene ID/Accession", "Lipid Probe", "Hit Annotation", "LogFC", "p value"))

### 2025-06-16 StudyInformation_includes/DCC_2025_include.qmd
Updating with data wrangling scripts and some suggested reviewer changes

<a href="/LipidProbe/N-Acylphosphatidylethanolamine.qmd" class="hover-preview">Bifunctional NAPE</a>
<span class="hover-image">
  <img src="/Structures/BF-NAPE.png" alt="Bifunctional NAPE" />
</span>

>>>>

<a href="/LipidProbe/N-Acylphosphatidylethanolamine.qmd" class="hover-preview">Bifunctional NAPE</a>
<span class="hover-image">
  <img src="/LipidProbe/Structures/BF-NAPE.png" alt="Bifunctional NAPE" />
</span>
